Continuous developmental change can explain discontinuities in word learning

### 2019-01-31 cogsci/cogsci.Rmd
second version cogsci

An unexpected outcome was that adult participants deviated slightly from the model's predicted numbers. While the model predicted recognition accuracy to be more sensitive to object distance when sound distance is high (and vice-versa), adult participants showed the opposite pattern. This deviation is interesting as it suggests that participants adapt their auditory-visual cue integration strategy to the ambiguity of the situation: They are more sensitive to the presence or absence of disambiguating visual information when sound ambiguity is higher (and vice-versa). \footnote{It is unclear whether the same pattern is found in children as the difference between accuracy scores is generally small compared to the uncertainty of the measurements.}

@@

One limitation of this work is that the model was fit to data from children at a relatively older age (4-5 years old) than what is typically studied in the literature (around 14 month-old). Nonetheless, this choice was related to two measurement difficulties. First, model fitting requires several, reliably-measured data points. In our case, data collection involved presenting participants with several trials across four conditions in a between-subject design. It would have been challenging to obtain these measures with infants. Second, the study of infants requires using specific testing methods. This fact would have made the quantitative study of development challenging. Indeed, it is not clear how data from these methods would directly compare to data collected using more typical methods in older children and adults.
<!--While an important conclusion of this paper is that the same process may apply to different developmental stages, it would be interesting to experimentally verify the generality of this conclusion in the context of early language acquisition. A challenge to this next step would be to construct a computational framework which would compare performance across heterogeneous testing methods (e.g., looking time vs. forced-choice). -->
In sum, this paper proposes a model that accounts for the development of an important aspect of word learning. The findings show that data can be explained based on a continuous process operating over similar representations across development. We used a case from word learning as an example, but the same idea might apply to other aspects of cognitive development that are typically thought of as stage-like (e.g., acquisition of a theory of mind). Computational models such as the one proposed here can help us investigate the extent to which such discontinuities emerge due to genuine qualitative changes, and the extent to which they reflect the granularity of the researchers' own measurement tools.

>>>>

An unexpected outcome was that adult participants deviated slightly from the model's numbers. While the model predicted  accuracy to be more sensitive to object distance when sound distance is higher (and vice-versa), adult participants showed the opposite pattern. This deviation is interesting as it suggests that participants are more likely to pay attention and integrate additional sources of information when ambiguity is higher. \footnote{It is unclear whether the same pattern is found in children as the difference between accuracy scores is generally small compared to the uncertainty of the measurements.}

@@

One limitation of this work is that the model was fit to data from children at a relatively older age (4-5 years old) than what is typically studied in the literature (around 14 month-old). Nonetheless, this choice was related to two measurement difficulties. First, model fitting requires several, reliably-measured data points. In our case, data collection involved presenting participants with several trials across four conditions in a between-subject design. It would have been challenging to obtain these measures with infants. Second, the study of infants requires using specific testing methods. This fact would have made the quantitative study of development challenging. Indeed, it is not clear how data from these methods would directly compare to data collected using more typical methods in older children and adults.
<!--While an important conclusion of this paper is that the same process may apply to different developmental stages, it would be interesting to experimentally verify the generality of this conclusion in the context of early language acquisition. A challenge to this next step would be to construct a computational framework which would compare performance across heterogeneous testing methods (e.g., looking time vs. forced-choice). -->

In sum, this paper proposes a model that accounts for the development of an important aspect of word learning. The findings show that data can be explained based on a continuous process operating over similar representations across development. We used a case from word learning as an example, but the same idea might apply to other aspects of cognitive development that are typically thought of as stage-like (e.g., acquisition of a theory of mind). Computational models such as the one proposed here can help us investigate the extent to which such discontinuities emerge due to genuine qualitative changes, and the extent to which they reflect the granularity of the researchers' own measurement tools.
